\mytitle

### 2020-05-14 paper.tex
speedups

Additional parameters that were used to calculate 
the velocities ($u_{id\infty}, u_{ijd}$) include: turbine height ($60$ meters),
ground roughness ($0.3$ meters), turbine rotor radius ($20$ meters), and a wind farm of size $2km \times 2km$\footnote{The implementation of constructing the 12 standard WFLO instances can be found in~\url{https://bit.ly/2YIjTl1}.}.

@@

For the DA, we used a single configuration that was found to work best across the different instances. 
The DA was run for increasingly large time limits ranging from $0.7$ seconds to $65$ seconds where we no longer observe an improvement in 
performance. The penalty terms $\lambda_1$ and $\lambda_2$ (see~Eq.~(\ref{QUBO}) are selected such that $\lambda_1= \lambda_2$ and sum of the two penalty terms, namely $\lambda_1+\lambda_2$, 
is set to the highest value that the Digital Annealer allows\footnote{The DA has an upper limit on the size of objective function coefficients.}.

@@

Recall that $\mathcal{W}$ is the set of our 12 standard WFLO instances. 
Let $b_w$ be the best energy solution (in terms of SS) obtained for an instance $w \in \mathcal{W}$ 
across the different methods.
%For example, the best solution (within 3600 seconds) for the second instance $w_2$ that aims at placing $30$ turbines on a $10\times10$ grid with a single wind regime is attained by QUBO-LS and the SS energy expression is $15742$ (kW); thus, we set $b_w = 15742$.
Then, for every point in time $t \in \{0,\ldots, 3600\}$ (measured in seconds), we can compute the relative error of a given approach per instance $w$. Let $\mathcal{A}$ be the set of solution approaches, $\mathcal{A}= \{QUBO-LS, QP-LS, DA-LS, ILP-LS, QP-SS\}$. Further, denote by $b_{w,t,a}$ the best solution attained before time $t$ of approach $a$ for instance $w$. Then, the mean relative error at time $t$ for approach $a$ on instance $w$ is given by \begin{equation}MRE(w,t,a) = \frac{b_w - b_{w,t,a}}{b_w}.\end{equation} Note that the expression is always positive, since we are solving a maximization problem.

We can now average over the different instances at time $t$. The average performance of approach $a$ at time $t$, $MRE(t,a)$ can be computed as, \begin{equation} MRE(t,a) = \frac{1}{|\mathcal{W}|}\sum_{ w\in\mathcal{W}} MRE(w,t,a).\end{equation}
We present $MRE(t,a)$ for each of the approaches to demonstrate the speed of convergence of the proposed LS-based methods.

@@

To demonstrate the speedup gained from using the DA, we compare the point in time when the DA reaches 
a competitive solution (10 seconds) to the point in time when QP-LS reaches the same solution in terms of energy.
Table~\ref{speedup} summarizes the performance of QP-LS when compared to that of the DA after 10 seconds.
We see a speedup factor of X. \todo{discuss with Chris}
%This is an encouraging result, especially due to the pessimistic computational complexity of WFLO.

@@

\subsection{Discussion and Practical Implications}

Our result show empirically that the quadratic programming models that we
introduced in Section~\ref{sec:QUBO4WFLO} (WFL-QCOP and WFL-QUBO) are useful in two ways.
First, the two problems are the basis for the QP-LS and QUBO-LS methods that solved with Gurobi. The two methods, namely QP-LS and QUBO-LS, turn out to be powerful approaches to solving WFLO. Specifically, QP-LS yields the best performance (after 3600 seconds) for most problem instances
and especially,
for the hard instances with 36 wind directions. In addition, the runtime performance of QP-LS is second
only to the Digital Annealer.  Second, the QUBO model can be solved with recent specialized hardware, Fujitsu's DA, as demonstrated by the DA-LS approach. DA-LS yields the best results in 7 out of 12 
instances, shows competitive performance on the other instances, and outperforms all methods in terms of runtime (\todo{speed-up ratio of ...})

>>>>

Additional parameters that were used to calculate 
the velocities ($u_{id\infty}, u_{ijd}$) include: turbine height ($60$ meters),
ground roughness ($0.3$ meters), turbine rotor radius ($20$ meters), and a wind farm of size $2km \times 2km$.\footnote{The implementation of constructing the 12 standard WFLO instances can be found in~\url{https://bit.ly/2YIjTl1}.}

@@

For the DA, we used a single configuration that was found to work best across the different instances. 
The DA was run for increasingly large time limits ranging from $0.7$ seconds to $65$ seconds where we no longer observe an improvement in 
performance. The penalty terms $\lambda_1$ and $\lambda_2$ (see~Eq.~(\ref{QUBO}) are selected such that $\lambda_1= \lambda_2$ and sum of the two penalty terms, namely $\lambda_1+\lambda_2$, 
is set to the highest value that the Digital Annealer allows.\footnote{The DA has an upper limit on the size of objective function coefficients.}

@@

Recall that $\mathcal{W}$ is the set of our 12 standard WFLO instances. 
Let $b_w$ be the best energy solution (in terms of SS) obtained for an instance $w \in \mathcal{W}$ 
across the different methods.
%For example, the best solution (within 3600 seconds) for the second instance $w_2$ that aims at placing $30$ turbines on a $10\times10$ grid with a single wind regime is attained by QUBO-LS and the SS energy expression is $15742$ (kW); thus, we set $b_w = 15742$.
Then, for every point in time $t \in \{0,\ldots, 3600\}$ (measured in seconds), we can compute the relative error of a given approach per instance $w$. Let $\mathcal{A}$ be the set of solution approaches, $\mathcal{A}= \{QUBO-LS, QP-LS, DA-LS, ILP-LS, QP-SS\}$. Further, denote by $b_{w,t,a}$ the best solution attained before time $t$ of approach $a$ for instance $w$. Then, the mean relative error at time $t$ for approach $a$ on instance $w$ is given by \begin{equation}MRE(w,t,a) = \frac{b_w - b_{w,t,a}}{b_w}.\end{equation} Note that the expression is always positive, since we are solving a maximization problem. We can now average over the different instances at time $t$. The average performance of approach $a$ at time $t$, $MRE(t,a)$ can be computed as, \begin{equation} MRE(t,a) = \frac{1}{|\mathcal{W}|}\sum_{ w\in\mathcal{W}} MRE(w,t,a).\end{equation}
We present $MRE(t,a)$ for each of the approaches to demonstrate the speed of convergence of the proposed LS-based methods.

\todo{@Chris - new part:} The last measure that we introduce is the speedup factor $S(a_1, a_2, t, w)$. The measure compares the time that it takes
approach $a_1 \in \mathcal{A}$ to
achieve the same energy level as the energy achieved by approach $a_2 \in \mathcal{A}$
within $t$ seconds when solving instance $w$ (energy levels are measured in terms of the sum-of-squares energy expression). 
Formally, $$S(a_1, a_2, t, w) = \frac{\min \big[\{ t' : E_{SS}(a_1, t',w) = E_{SS}(a_2, t, w) \} \cup \{3600\} \big] }{t}, $$
with $E_{SS}(a, t, w)$ being the SS energy value under approach $a$ for instance $w$ by time $t$. The speedup factor 
is useful to measure how many times faster (or slower) approach $a_1$ solves problem $w$ compared to approach $a_2$.   
For example, $$S(QP-LS, DA-LS, 10sec, w_1)$$ is computed as the ratio between the time that it takes QP-LS to 
reach the same quality of solution as DA-LS achieves after $10$ seconds (over problem $w_1$). Since
we limit the runtime of the exact methods by $3600$ seconds, we may get a lower bound on the speedup factor. 
Therefore, if for the example above the result is $3600/10$, we shall interpret it as a speedup factor of \emph{at least} $360$. \todo{@Chris - new part ends here.}

@@

\todo{@Chris - new part:} To demonstrate the gain in runtime from using the DA, we compare the average speedup factor between
the QP-LS (the dominant exact method) to a DA-LS solutions after $1$ and $10$ seconds.
In other words, $w \in \mathcal{W}$ we computed $S(QP-LS, DA-LS, 1sec, w)$ and $S(QP-LS, DA-LS, 10sec, w)$
for every $w$ 
and averaged the speedup factors across the instances.

As a result, the average speedup factors were found to be $666$ and $100$ for $1$ and $10$ seconds, respectively.\footnote{The speedup factors  were rounded to have 0 values after the floating point.} These results show that the DA achieves competitive solutions after running for $1$ second.
Note that since for many of the instances QP-LS did not attain
the energy level of DA-LS at time $t$ (1 or 10 seconds), we consider the speedup factor to be a lower bound, meaning
that the DA-LS gives potentially an even higher speedup factor, which we did not compute due to the time limit of $3600$. \todo{@Chris - new part ends here}
%This is an encouraging result, especially due to the pessimistic computational complexity of WFLO.

@@

\subsection{Discussion and Practical Implications}
\todo{@Chris - please check discussion:}
Our result show empirically that the quadratic programming models that we
introduced in Section~\ref{sec:QUBO4WFLO} (WFL-QCOP and WFL-QUBO) are useful in two ways.
First, the two problems are the basis for the QP-LS and QUBO-LS methods that solved with Gurobi. The two methods, namely QP-LS and QUBO-LS, turn out to be powerful approaches to solving WFLO. Specifically, QP-LS yields the best performance (after 3600 seconds) for most problem instances
and especially,
for the hard instances with 36 wind directions. In addition, the runtime performance of QP-LS is second
only to the Digital Annealer.  Second, the QUBO model can be solved with recent specialized hardware, Fujitsu's DA, as demonstrated by the DA-LS approach. DA-LS yields the best results in 7 out of 12 
instances, shows competitive performance on the other instances, and outperforms all methods in terms of runtime (speedup factor of 2 orders of magnitude when the DA runs for $1$ second).
